---
title: "Police Shootings EDA"
format: html
editor: visual
---

```{r setup}
#| include: false
pacman::p_load(rio, tidyverse)
police_shootings <- read_csv("data/police_shootings.csv")
```

# Goal

-   Exploratory Data Analysis of Police Shootings.

# Data

```{r}
police_shootings
```

-   Grain: Police shooting events

Nulls (Bar Plot, Time Series)

```{r}
police_shootings |>
  mutate_all(is.na) |>
  summarise_all(sum) |>
  pivot_longer(cols = names(police_shootings)) |>
  mutate(p = round(100*value/9345, 1))
```

# Victim Demographics

## Age (Bar Plot, Time Series)

-   Discussion

```{r}
# Plot/Graph
police_shootings |>
  ggplot(aes(y = age)) +
  geom_histogram(binwidth = 10, fill = "darkblue") +
  ylim(90, 0)
```

## Gender (Bar Plot, Time Series)

-   Discussion

```{r}
# Table
ps_gender <-
  police_shootings |>
  filter(!is.na(gender)) |>
  count(gender) |>
  mutate(p = round(100*n/sum(n), 2))

ps_year_gender <-
  police_shootings |>
  filter(!is.na(gender)) |>
  group_by(year, gender) |>
  summarize(n = n()) |>
  group_by(year) |>
  mutate(p = round(100*n/sum(n), 2))

ps_gender
ps_year_gender
```

```{r}
# Plot/Graph
ps_gender |>
  ggplot(aes(x = gender, y = p)) +
  geom_col()

ps_year_gender |>
  ggplot(aes(x = year, y = p, color = gender)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(
    breaks = unique(ps_year_gender$year)
  )
```

## Race

-   Discussion

```{r}
# Table

```

```{r}
# Plot/Graph

```

## Race Source (Bar Plot, Time Series)

```{r}
# Table

```

```{r}
# Plot/Graph

```

## Was Mental Illness Related (Bar Plot, Time Series)

-   Discussion

```{r}
# Table

```

```{r}
# Plot/Graph

```

# Event Specifics

## Threat Type (Bar Plot, Time Series)

-   Discussion

```{r}
# Table

```

```{r}
# Plot/Graph

```

## Flee Status (Bar Plot, Time Series)

-   Discussion

```{r}
# Table

```

```{r}
# Plot/Graph

```

## Armed With (Bar Plot, Time Series)

-   Discussion

```{r}
# Table

```

```{r}
# Plot/Graph

```

## Location Precision (Bar Plot, Time Series)

-   Discussion

```{r}
# Together
ps_state <-
  police_shootings |> 
  count(state) |>
  mutate(p = round(100*n/sum(n),2)) |>
  arrange(desc(p))
ps_state
```

```{r}
pacman::p_load(janitor, sf)
shapefile <-
  read_sf("data/tl_2022_us_state.shp") |>
  clean_names()

shapefile |>
  filter(region %in% 1:4, !name %in% c("Alaska", "Hawaii")) |>
  inner_join(ps_state, by = c("stusps" = "state")) |>
  ggplot() +
  geom_sf(aes(fill = n, geometry = geometry))
```

## Body Camera (Bar Plot, Time Series)

-   Discussion

```{r}
# Table

```

```{r}
# Plot/Graph

```

# Big Questions

## Does the total number of police shootings increase every year?

-   Discussion

```{r}

```

## Which demographic groups are most impacted?

-   Discussion

```{r}

```

## Surprise

Other than the sheer number of police shootings, what pattern did you see in this data which most surprised you or caused you to ask additional questions?

For example, was there a change in any of the trend lines which surprised you, or a demographic or event group which seemed larger or smaller than you expected?

Use one or more tables/plots to support your answer.

-   Discussion

```{r}

```

# Heat Maps

## Race x Body Camera

```{r}
police_shootings |>
  count(race, body_camera) |>
  group_by(race) |>
  mutate(p = 100*n/sum(n)) |>
  filter(race !="Multiracial") |>
  ggplot(aes(x = body_camera, y = race, fill = p)) +
  geom_tile()
```

## Race x Flee Status

```{r}

```
