---
title: "Police Shootings EDA"
format: html
editor: visual
---

```{r setup}
#| include: false
pacman::p_load(rio, tidyverse)
police_shootings <- read_csv("data/police_shootings.csv")
```

# Goal

-   Exploratory Data Analysis of Police Shootings.

# Data

```{r}
police_shootings
```

-   Grain: Police shooting events

Nulls (Bar Plot, Time Series)

```{r}
police_shootings |>
  mutate_all(is.na) |>
  summarise_all(sum) |>
  pivot_longer(cols = names(police_shootings)) |>
  mutate(p = round(100*value/9345, 1))
```

# Victim Demographics

## Age (Bar Plot, Time Series)

-   Discussion

```{r}
# Plot/Graph
police_shootings |>
  ggplot(aes(y = age)) +
  geom_histogram(binwidth = 10, fill = "darkblue") +
  ylim(90, 0)
```

## Gender (Bar Plot, Time Series)

-   Discussion

```{r}
# Table
ps_gender <-
  police_shootings |>
  filter(!is.na(gender)) |>
  count(gender) |>
  mutate(p = round(100*n/sum(n), 2))

ps_year_gender <-
  police_shootings |>
  filter(!is.na(gender)) |>
  group_by(year, gender) |>
  summarize(n = n()) |>
  group_by(year) |>
  mutate(p = round(100*n/sum(n), 2))

ps_gender
ps_year_gender
```

```{r}
# Plot/Graph
ps_gender |>
  ggplot(aes(x = gender, y = p)) +
  geom_col()

ps_year_gender |>
  ggplot(aes(x = year, y = p, color = gender)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(
    breaks = unique(ps_year_gender$year)
  )
```

## Race

-   Discussion

```{r}
# Table
ps_race <-
  police_shootings |>
  mutate(race = if_else(is.na(race), "Unknown", race)) |>
  filter(!str_detect(race, ";")) |>
  count(race) |>
  mutate(p = round(100*n/sum(n), 2))

ps_year_race <-
  police_shootings |>
  filter(!is.na(race), !str_detect(race, ";")) |>
  group_by(year, race) |>
  summarize(n = n()) |>
  group_by(year) |>
  mutate(p = round(100*n/sum(n), 2))

ps_race
ps_year_race
```

```{r}
# Plot/Graph
ps_race |>
  ggplot(aes(x = race, y = p)) +
  geom_col()

ps_year_race |>
  ggplot(aes(x = year, y = p, color = race)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(
    breaks = unique(ps_year_race$year)
  )
```

## Race Source (Bar Plot, Time Series)

```{r}
# Table
ps_race_source <-
  police_shootings |>
  filter(!is.na(race_source)) |>
  count(race_source) |>
  mutate(p = round(100*n/sum(n), 2))

ps_year_race_source <-
  police_shootings |>
  filter(!is.na(race_source)) |>
  group_by(year, race_source) |>
  summarize(n = n()) |>
  group_by(year) |>
  mutate(p = round(100*n/sum(n), 2))

ps_race_source
ps_year_race_source
```

```{r}
# Plot/Graph
ps_race_source |>
  ggplot(aes(x = race_source, y = p)) +
  geom_col()

ps_year_race_source |>
  ggplot(aes(x = year, y = p, color = race_source)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(
    breaks = unique(ps_year_race_source$year)
  )
```

## Was Mental Illness Related (Bar Plot, Time Series)

-   Discussion

```{r}
# Table
ps_was_mental_illness_related <-
  police_shootings |>
  filter(!is.na(was_mental_illness_related)) |>
  count(was_mental_illness_related) |>
  mutate(p = round(100*n/sum(n), 2))

ps_year_was_mental_illness_related <-
  police_shootings |>
  filter(!is.na(was_mental_illness_related)) |>
  group_by(year, was_mental_illness_related) |>
  summarize(n = n()) |>
  group_by(year) |>
  mutate(p = round(100*n/sum(n), 2))

ps_was_mental_illness_related
ps_year_was_mental_illness_related
```

```{r}
# Plot/Graph
ps_was_mental_illness_related |>
  ggplot(aes(x = was_mental_illness_related, y = p)) +
  geom_col()

ps_year_was_mental_illness_related |>
  ggplot(aes(x = year, y = p, color = was_mental_illness_related)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(
    breaks = unique(ps_year_was_mental_illness_related$year)
  )
```

# Event Specifics

## Threat Type (Bar Plot, Time Series)

-   Discussion

```{r}
# Table
ps_threat_type <-
  police_shootings |>
  filter(!is.na(threat_type)) |>
  count(threat_type) |>
  mutate(p = round(100*n/sum(n), 2))

ps_year_threat_type <-
  police_shootings |>
  filter(!is.na(threat_type)) |>
  group_by(year, threat_type) |>
  summarize(n = n()) |>
  group_by(year) |>
  mutate(p = round(100*n/sum(n), 2))

ps_threat_type
ps_year_threat_type
```

```{r}
# Plot/Graph
ps_threat_type |>
  ggplot(aes(x = threat_type, y = p)) +
  geom_col()

ps_year_threat_type |>
  ggplot(aes(x = year, y = p, color = threat_type)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(
    breaks = unique(ps_year_threat_type$year)
  )
```

## Flee Status (Bar Plot, Time Series)

-   Discussion

```{r}
# Table
ps_flee_status <-
  police_shootings |>
  filter(!is.na(flee_status)) |>
  count(flee_status) |>
  mutate(p = round(100*n/sum(n), 2))

ps_year_flee_status <-
  police_shootings |>
  filter(!is.na(flee_status)) |>
  group_by(year, flee_status) |>
  summarize(n = n()) |>
  group_by(year) |>
  mutate(p = round(100*n/sum(n), 2))

ps_flee_status
ps_year_flee_status
```

```{r}
# Plot/Graph
ps_flee_status |>
  ggplot(aes(x = flee_status, y = p)) +
  geom_col()

ps_year_flee_status |>
  ggplot(aes(x = year, y = p, color = flee_status)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(
    breaks = unique(ps_year_flee_status$year)
  )
```

## Armed With (Bar Plot, Time Series)

-   Discussion

```{r}
# Table
ps_armed_with <-
  police_shootings |>
  filter(!is.na(armed_with)) |>
  count(armed_with) |>
  mutate(p = round(100*n/sum(n), 2))

ps_year_armed_with <-
  police_shootings |>
  filter(!is.na(armed_with)) |>
  group_by(year, armed_with) |>
  summarize(n = n()) |>
  group_by(year) |>
  mutate(p = round(100*n/sum(n), 2))

ps_armed_with
ps_year_armed_with
```

```{r}
# Plot/Graph
ps_armed_with |>
  ggplot(aes(x = armed_with, y = p)) +
  geom_col()

ps_year_armed_with |>
  ggplot(aes(x = year, y = p, color = armed_with)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(
    breaks = unique(ps_year_armed_with$year)
  )
```

## Location Precision (Bar Plot, Time Series)

-   Discussion

```{r}
# Together
ps_state <-
  police_shootings |> 
  count(state) |>
  mutate(p = round(100*n/sum(n),2)) |>
  arrange(desc(p))
ps_state
```

```{r}
pacman::p_load(janitor, sf)
shapefile <-
  read_sf("data/tl_2022_us_state.shp") |>
  clean_names()

shapefile |>
  filter(region %in% 1:4, !name %in% c("Alaska", "Hawaii")) |>
  inner_join(ps_state, by = c("stusps" = "state")) |>
  ggplot() +
  geom_sf(aes(fill = n, geometry = geometry))
```

## Body Camera (Bar Plot, Time Series)

-   Discussion

```{r}
# Table
ps_body_camera <-
  police_shootings |>
  filter(!is.na(body_camera)) |>
  count(body_camera) |>
  mutate(p = round(100*n/sum(n), 2))

ps_year_body_camera <-
  police_shootings |>
  filter(!is.na(body_camera)) |>
  group_by(year, body_camera) |>
  summarize(n = n()) |>
  group_by(year) |>
  mutate(p = round(100*n/sum(n), 2))

ps_body_camera
ps_year_body_camera
```

```{r}
# Plot/Graph
ps_body_camera |>
  ggplot(aes(x = body_camera, y = p)) +
  geom_col()

ps_year_body_camera |>
  ggplot(aes(x = year, y = p, color = body_camera)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(
    breaks = unique(ps_year_body_camera$year)
  )
```

# Big Questions

## Does the total number of police shootings increase every year?

-   Discussion

```{r}

```

## Which demographic groups are most impacted?

-   Discussion

```{r}

```

## Surprise

Other than the sheer number of police shootings, what pattern did you see in this data which most surprised you or caused you to ask additional questions?

For example, was there a change in any of the trend lines which surprised you, or a demographic or event group which seemed larger or smaller than you expected?

Use one or more tables/plots to support your answer.

-   Discussion

```{r}

```


# Heat Maps

## Race x Body Camera

```{r}
police_shootings |>
  count(race, body_camera) |>
  group_by(race) |>
  mutate(p = 100*n/sum(n)) |>
  filter(race !="Multiracial") |>
  ggplot(aes(x = body_camera, y = race, fill = p)) +
  geom_tile()
```



## Race x Flee Status

```{r}

```
